Updated Constraints on Self-Interacting Dark Matter from Supernova 1987A

### 2017-04-20 draft.tex
Raffelt as supernova property citation?

The quantity of interest is the rate of energy emission through dark gauge bosons. 
From the spin-summed, squared amplitudes described in the previous subsection, 
the energy emission rate is obtained by integrating over the phase space, 
and adding a factor of the energy of the emitted particle. To be specific, the 
energy emission rate per unit volume is 
%
\begin{widetext}
\beq
\label{eq:rate1}
Q_i = (2\pi)^4 \int\, E_{\mathrm{A}'} \, \sum_{s_1,s_2}\, \vert \mathcal{M}_i \vert^2 f(p_1) f(p_2)\delta^{(4)}(p_1+p_2-p_3-p_4-\qa)\, \dd \Pi,
\eeq
\end{widetext}
%
where 
%
\begin{equation}
d\Pi = \frac{\dd^3 \vec{q}_{\mathrm{A}'}}{(2\pi)^3 2E_{\mathrm{A}'}}\, \prod_{j=1}^{4}\, \frac{\dd^3 \vec{p}_j}{(2\pi^3) 2E_j}
\end{equation} 
is the Lorentz-invariant phase space interval, 
$ E_{\mathrm{A}'}$ is the energy of the emitted $\Aprime$ boson, 
$f(p)$ are the phase-space densities of the incoming nucleons, and 
the index $i$ on $Q_i$ refers to either the pp or pn processes. 
The nucleons in the core are comfortably non-degenerate and non-relativistic, 
so the Pauli blocking factor is omitted from Eq.~(\ref{eq:rate1}) and we take 
all nucleons to have a Maxwell-Boltzmann phase-space distribution distribution, 
\beq
f(p) =  \frac{n_b}{2} \left( \frac{2 \pi}{M_N T} \right)^{3/2} \exp \left(-\frac{p^2}{2 M_N T} \right).
\eeq
%
We choose a baryon number density of $n_{\mathrm{b}} \approx 1.8 \times 10^{38}\, \mathrm{cm}^{-3}$ and a 
core supernova temperature of $T = 30\, \mathrm{MeV}$, both of which are typical choices and thought to 
be representative of supernova cores. \cite{snova_core} \arz{Need to put a citation here for the properties of supernova cores. 
I can probably find something in the literature, but please put something here if you have a citation.} \cbm{That okay?}

@@

The second effect that produces an upper bound on the excluded region 
comes from considering trapping of dark bosons within the supernova. 
With a large enough coupling, the dark photons will thermalize and then 
will be emitted from an approximately spherical ``dark photosphere" at the radial position 
where the $\Aprime$ mean free path becomes larger than the typical size of the supernova core. 
In this case the luminosity is given simply by Stefan's law, 
\beq
L_{\mathrm{A'}}  = 4\pi \rdp^2 T_{\mathrm{A'}}^4 \sigma,
\eeq
where $\rdp$ is now the radius of the emitting shell and $T_{\mathrm{A'}}$ its temperature. 
We estimate the radius of this dark photosphere as $\rdp=10$ km, 
because the density of the supernova drops drastically around that point. \cite{snova_density} 
\arz{We need a citation for this statement, but I do not know one off of the top of my head.}\cbm{And that?}
The bound on the luminosity can then be recast as a bound on $T_{\mathrm{A}'}$, 
%
\beq
T_{\mathrm{A}'} \lesssim 9.6\rm\ MeV.
\eeq
%
That bound can then be translated into the desired bound on the coupling as a function of mass by adopting a 
simple model for the supernova atmosphere, assuming that the particles are emitted from the dark photosphere 
at a point where the optical depth to scattering is $\tau = 2/3 $, and finding the temperature that corresponds 
to that optical depth.

>>>>

The quantity of interest is the rate of energy emission through dark gauge bosons. 
From the spin-summed, squared amplitudes described in the previous subsection, 
the energy emission rate is obtained by integrating over the phase space, 
and adding a factor of the energy of the emitted particle. To be specific, the 
energy emission rate per unit volume is 
%
\begin{widetext}
\beq
\label{eq:rate1}
Q_i = (2\pi)^4 \int\, E_{\mathrm{A}'} \, \sum_{s_1,s_2}\, \vert \mathcal{M}_i \vert^2 f(p_1) f(p_2)\delta^{(4)}(p_1+p_2-p_3-p_4-\qa)\, \dd \Pi,
\eeq
\end{widetext}
%
where 
%
\begin{equation}
d\Pi = \frac{\dd^3 \vec{q}_{\mathrm{A}'}}{(2\pi)^3 2E_{\mathrm{A}'}}\, \prod_{j=1}^{4}\, \frac{\dd^3 \vec{p}_j}{(2\pi^3) 2E_j}
\end{equation} 
is the Lorentz-invariant phase space interval, 
$ E_{\mathrm{A}'}$ is the energy of the emitted $\Aprime$ boson, 
$f(p)$ are the phase-space densities of the incoming nucleons, and 
the index $i$ on $Q_i$ refers to either the pp or pn processes. 
The nucleons in the core are comfortably non-degenerate and non-relativistic, 
so the Pauli blocking factor is omitted from Eq.~(\ref{eq:rate1}) and we take 
all nucleons to have a Maxwell-Boltzmann phase-space distribution distribution, 
\beq
f(p) =  \frac{n_b}{2} \left( \frac{2 \pi}{M_N T} \right)^{3/2} \exp \left(-\frac{p^2}{2 M_N T} \right).
\eeq
%
We choose a baryon number density of $n_{\mathrm{b}} \approx 1.8 \times 10^{38}\, \mathrm{cm}^{-3}$ and a 
core supernova temperature of $T = 30\, \mathrm{MeV}$, both of which are typical choices and thought to 
be representative of supernova cores. \arz{Need to put a citation here for the properties of supernova cores. 
I can probably find something in the literature, but please put something here if you have a citation.} \cbm{On second thought, is Raffelt not good enough for both of these? That's where everybody started}

@@

The second effect that produces an upper bound on the excluded region 
comes from considering trapping of dark bosons within the supernova. 
With a large enough coupling, the dark photons will thermalize and then 
will be emitted from an approximately spherical ``dark photosphere" at the radial position 
where the $\Aprime$ mean free path becomes larger than the typical size of the supernova core. 
In this case the luminosity is given simply by Stefan's law, 
\beq
L_{\mathrm{A'}}  = 4\pi \rdp^2 T_{\mathrm{A'}}^4 \sigma,
\eeq
where $\rdp$ is now the radius of the emitting shell and $T_{\mathrm{A'}}$ its temperature. 
We estimate the radius of this dark photosphere as $\rdp=10$ km, 
because the density of the supernova drops drastically around that point. 
\arz{We need a citation for this statement, but I do not know one off of the top of my head.}
The bound on the luminosity can then be recast as a bound on $T_{\mathrm{A}'}$, 
%
\beq
T_{\mathrm{A}'} \lesssim 9.6\rm\ MeV.
\eeq
%
That bound can then be translated into the desired bound on the coupling as a function of mass by adopting a 
simple model for the supernova atmosphere, assuming that the particles are emitted from the dark photosphere 
at a point where the optical depth to scattering is $\tau = 2/3 $, and finding the temperature that corresponds 
to that optical depth.
